HO unification from object language to meta language

### 2024-03-21 paper/appendix.tex
renaming

  type move assignment -> list tm -> tm -> o.
  move (abs Bo)        [H|L] R :- move (Bo H) L R.
  move (val A)         []    A :- !.
  move (val (uva N A)) L     (uva N X) :- std.append A L X.

  type beta tm -> list tm -> tm -> o.
  beta A [] A.
  beta (lam Bo) [H | L] R :- beta (Bo H) L R.
  beta (app A) L (app X) :- std.append A L X.
  beta (uva N A) L (uva N A') :- std.append A L A'.
  beta (con H) L (app [con H | L]).

@@

  type locale ho.tm -> o.
  type get-scope-aux ho.tm -> list ho.tm -> o.
  get-scope-aux (ho.con _) [].
  get-scope-aux (ho.uva _ L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux (ho.lam B) L1 :- 
    pi x\ locale x => get-scope-aux (B x) L1.
  get-scope-aux (ho.app L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux X [X] :- name X, not (locale X).
  get-scope-aux X [] :- name X, (locale X).

@@

  type comp fo.fm -> ho.tm -> mappings -> mappings -> links -> 
    links -> ho.subst -> ho.subst -> o.
  comp (fo.fcon X) (ho.con X)        Map1 Map1 L1 L1 S1 S1.
  comp (fo.flam F) (ho.uva E0 Scope) Map1 Map2 L1 L3 S1 S3 :-
    (pi x\ maybe-eta x (F x) [x]), !,
      mem.new S1 E0 S2,
      comp-lam F F1 Map1 Map2 L1 L2 S2 S3,
      (pi x\ locale x => get-scope (F1 x) Scope), 
      NewLink = ho.val(link-eta (ho.uva E0 Scope) (ho.lam F1)), 
      L3 = [NewLink| L2].
  comp (fo.flam F) (ho.lam F1)       Map1 Map2 L1 L2 S1 S2 :- 
    comp-lam F F1 Map1 Map2 L1 L2 S1 S2.
  comp (fo.fuva N) (ho.uva M []) Map1 Map2 L L S S1 :- 
    alloc Map1 Map2 (fv N) (hv M (arity z)) S S1.
  comp (fo.fapp [fo.fuva N|Ag]) (ho.uva M Scope) Map1 Map2 L1 L1 S1 S2 :-
    pattern-fragment Ag, !,
      fold6 comp Ag Scope Map1 Map1 L1 L1 S1 S1,
      len Ag Arity, 
      alloc Map1 Map2 (fv N) (hv M (arity Arity)) S1 S2.
  comp (fo.fapp [fo.fuva N|Ag]) (ho.uva M' Scope) Map1 Map3 L1 L3 S1 S4 :- !,
    pattern-fragment-prefix Ag PF NPF,
      fold6 comp PF PF1 Map1 Map1 L1 L1 S1 S1,
      fold6 comp NPF NPF1 Map1 Map2 L1 L2 S1 S2,
      len PF Arity, 
      alloc Map2 Map3 (fv N) (hv M (arity Arity)) S2 S3,
      get-scope (ho.app {std.append PF1 NPF1}) Scope, 
      mem.new S3 M' S4,
      L3 = [@val-link-beta (ho.uva M' Scope) (ho.app[ho.uva M PF1 | NPF1]) | L2].
  comp (fo.fapp A) (ho.app A1) Map1 Map2 L1 L2 S1 S2 :- 
    fold6 comp A A1 Map1 Map2 L1 L2 S1 S2.

@@

  type deduplicate-mappings mappings -> mappings -> 
          ho.subst -> ho.subst -> links -> links -> o.
  deduplicate-mappings [] [] H H L L.
  deduplicate-mappings [(mapping (fv O) (hv M (arity LenM)) as X1) | Map1] Map2 H1 H3 L1 L3 :-
    take-list Map1 (mapping (fv O) (hv M' (arity LenM'))) _, !,
    make-eta-link LenM LenM' M M' [] New H1 H2,
    std.append New L1 L2,
    deduplicate-mappings Map1 Map2 H2 H3 L2 L3.
  deduplicate-mappings [A|As] [A|Bs] H1 H2 L1 L2 :- 
    deduplicate-mappings As Bs H1 H2 L1 L2, !.
\end{elpicode}

@@

  type contract-rigid list ho.tm -> ho.tm -> ho.tm -> o.
  contract-rigid L (ho.lam F) T :- 
    pi x\ contract-rigid [x|L] (F x) T. % also checks H Prefix does not see x
  contract-rigid L (ho.app [H|Args]) T :- 
    std.rev L LRev, std.appendR Prefix LRev Args,
    if (Prefix = []) (T = H) (T = ho.app [H|Prefix]).

@@

  progress-beta-link (ho.uva _ _ as T) (ho.app[ho.uva Ad1 Scope1 | L1] as T1) S1 S3 NewLinks :-
    std.append Scope1 L1 Scope1L,
    pattern-fragment-prefix Scope1L Scope2 L2,
    not (Scope1 = Scope2), !,
    mem.new S1 Ad2 S2,
    len Scope1 Scope1Len,
    len Scope2 Scope2Len,
    make-eta-link Scope1Len Scope2Len Ad1 Ad2 [] LinkEta S2 S3,
    if (L2 = []) (NewLinks = LinkEta, T2 = ho.uva Ad2 Scope2) 
      (T2 = ho.app [ho.uva Ad2 Scope2 | L2], 
      NewLinks = [@val-link-beta T T2 | LinkEta]).

@@

  type solve-links links -> links -> ho.subst -> ho.subst -> o.
  solve-links [] [] X X.
  solve-links [A|L1] [A|L3] S S2 :- take-link A L1 B L2, !,
    same-link-eta A B S S1, 
    solve-links L2 L3 S1 S2.
  solve-links [L0|L1] L3 S S2 :- deref-link S L0 L,
    solve-link-abs L R S S1, !,
    solve-links L1 L2 S1 S2, std.append R L2 L3.
\end{elpicode}

@@

  type complete-mapping ho.subst -> ho.subst ->   
    mappings -> mappings -> fo.subst -> fo.subst -> o.
  complete-mapping _ [] L L F F.
  complete-mapping H [none | Tl] L1 L2 F1 F2 :-   
    complete-mapping H Tl L1 L2 F1 F2.
  complete-mapping H [some T0 | Tl] L1 L3 F1 F3 :-
    ho.deref-assmt H T0 T,
    complete-mapping-under-ass H T L1 L2 F1 F2, 
    std.append L1 L2 LAll,
    complete-mapping H Tl LAll L3 F2 F3.

>>>>

  type move assignment -> list tm -> tm -> o.
  move (abs Bo)        [H|L] R :- move (Bo H) L R.
  move (val A)         []    A :- !.
  move (val (uva N A)) L     (uva N X) :- append A L X.

  type beta tm -> list tm -> tm -> o.
  beta A [] A.
  beta (lam Bo) [H | L] R :- beta (Bo H) L R.
  beta (app A) L (app X) :- append A L X.
  beta (uva N A) L (uva N A') :- append A L A'.
  beta (con H) L (app [con H | L]).

@@

  type locally-bound ho.tm -> o.
  type get-scope-aux ho.tm -> list ho.tm -> o.
  get-scope-aux (ho.con _) [].
  get-scope-aux (ho.uva _ L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux (ho.lam B) L1 :- 
    pi x\ locally-bound x => get-scope-aux (B x) L1.
  get-scope-aux (ho.app L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux X [X] :- name X, not (locally-bound X).
  get-scope-aux X [] :- name X, (locally-bound X).

@@

  type comp fo.fm -> ho.tm -> mappings -> mappings -> links -> 
    links -> ho.subst -> ho.subst -> o.
  comp (fo.fcon X) (ho.con X)        Map1 Map1 L1 L1 S1 S1.
  comp (fo.flam F) (ho.uva E0 Scope) Map1 Map2 L1 L3 S1 S3 :-
    (pi x\ maybe-eta x (F x) [x]), !,
      mem.new S1 E0 S2,
      comp-lam F F1 Map1 Map2 L1 L2 S2 S3,
      (pi x\ locally-bound x => get-scope (F1 x) Scope), 
      NewLink = ho.val(link-eta (ho.uva E0 Scope) (ho.lam F1)), 
      L3 = [NewLink| L2].
  comp (fo.flam F) (ho.lam F1)       Map1 Map2 L1 L2 S1 S2 :- 
    comp-lam F F1 Map1 Map2 L1 L2 S1 S2.
  comp (fo.fuva N) (ho.uva M []) Map1 Map2 L L S S1 :- 
    alloc Map1 Map2 (fv N) (hv M (arity z)) S S1.
  comp (fo.fapp [fo.fuva N|Ag]) (ho.uva M Scope) Map1 Map2 L1 L1 S1 S2 :-
    pattern-fragment Ag, !,
      fold6 comp Ag Scope Map1 Map1 L1 L1 S1 S1,
      len Ag Arity, 
      alloc Map1 Map2 (fv N) (hv M (arity Arity)) S1 S2.
  comp (fo.fapp [fo.fuva N|Ag]) (ho.uva M' Scope) Map1 Map3 L1 L3 S1 S4 :- !,
    pattern-fragment-prefix Ag PF NPF,
      fold6 comp PF PF1 Map1 Map1 L1 L1 S1 S1,
      fold6 comp NPF NPF1 Map1 Map2 L1 L2 S1 S2,
      len PF Arity, 
      alloc Map2 Map3 (fv N) (hv M (arity Arity)) S2 S3,
      get-scope (ho.app {append PF1 NPF1}) Scope, 
      mem.new S3 M' S4,
      L3 = [@val-link-beta (ho.uva M' Scope) (ho.app[ho.uva M PF1 | NPF1]) | L2].
  comp (fo.fapp A) (ho.app A1) Map1 Map2 L1 L2 S1 S2 :- 
    fold6 comp A A1 Map1 Map2 L1 L2 S1 S2.

@@

  type deduplicate-mappings mappings -> mappings -> 
          ho.subst -> ho.subst -> links -> links -> o.
  deduplicate-mappings [] [] H H L L.
  deduplicate-mappings [(mapping (fv O) (hv M (arity LenM)) as X1) | Map1] Map2 H1 H3 L1 L3 :-
    take-list Map1 (mapping (fv O) (hv M' (arity LenM'))) _, !,
    make-eta-link LenM LenM' M M' [] New H1 H2,
    append New L1 L2,
    deduplicate-mappings Map1 Map2 H2 H3 L2 L3.
  deduplicate-mappings [A|As] [A|Bs] H1 H2 L1 L2 :- 
    deduplicate-mappings As Bs H1 H2 L1 L2, !.
\end{elpicode}

@@

  type contract-rigid list ho.tm -> ho.tm -> ho.tm -> o.
  contract-rigid L (ho.lam F) T :- 
    pi x\ contract-rigid [x|L] (F x) T. % also checks H Prefix does not see x
  contract-rigid L (ho.app [H|Args]) T :- 
    std.rev L LRev, append Prefix LRev Args,
    if (Prefix = []) (T = H) (T = ho.app [H|Prefix]).

@@

  progress-beta-link (ho.uva _ _ as T) (ho.app[ho.uva Ad1 Scope1 | L1] as T1) S1 S3 NewLinks :-
    append Scope1 L1 Scope1L,
    pattern-fragment-prefix Scope1L Scope2 L2,
    not (Scope1 = Scope2), !,
    mem.new S1 Ad2 S2,
    len Scope1 Scope1Len,
    len Scope2 Scope2Len,
    make-eta-link Scope1Len Scope2Len Ad1 Ad2 [] LinkEta S2 S3,
    if (L2 = []) (NewLinks = LinkEta, T2 = ho.uva Ad2 Scope2) 
      (T2 = ho.app [ho.uva Ad2 Scope2 | L2], 
      NewLinks = [@val-link-beta T T2 | LinkEta]).

@@

  type solve-links links -> links -> ho.subst -> ho.subst -> o.
  solve-links [] [] X X.
  solve-links [A|L1] [A|L3] S S2 :- take-link A L1 B L2, !,
    same-link-eta A B S S1, 
    solve-links L2 L3 S1 S2.
  solve-links [L0|L1] L3 S S2 :- deref-link S L0 L,
    solve-link-abs L R S S1, !,
    solve-links L1 L2 S1 S2, append R L2 L3.
\end{elpicode}

@@

  type complete-mapping ho.subst -> ho.subst ->   
    mappings -> mappings -> fo.subst -> fo.subst -> o.
  complete-mapping _ [] L L F F.
  complete-mapping H [none | Tl] L1 L2 F1 F2 :-   
    complete-mapping H Tl L1 L2 F1 F2.
  complete-mapping H [some T0 | Tl] L1 L3 F1 F3 :-
    ho.deref-assmt H T0 T,
    complete-mapping-under-ass H T L1 L2 F1 F2, 
    append L1 L2 LAll,
    complete-mapping H Tl LAll L3 F2 F3.
